Stats Project 2

### 2022-12-12 article.Rmd
updated final render of PDF + images

---
title: "Stats Project 2"
author: Abhay Sharma | AbhaySh4r.github.io
output:
  pdf_document: default
  html_notebook: default
---
```{r include=FALSE}
library(tidyverse)
library(plotly)
library(car)
library(caret)
library(olsrr)
library(corrplot)
```

@@

@@

For this project, we are provided a housing data set from the website Kaggle that contains a large collection of parameters that are likely to be correlated with the Sale Price of their respective houses. Within this data set we find 79 parameters, some of which must ultimately be used to build a predictive model to extrapolate housing prices given a certain set of these parameters.

@@

\subsection{Model Generation and Analysis}
To generate an initial model, we will look at a rudimentary approach to linear regression to see what our data does;

```{r echo = TRUE, results = 'hide', warning=FALSE}
housing = read.csv(file = "train.csv")
numeric_housing = housing %>% select_if(is.numeric)
a1_set = housing %>% select(SalePrice, Neighborhood, GrLivArea) %>% filter( Neighborhood == c("NAmes", "Edwards", "BrkSide"))
fit <- lm(SalePrice ~ GrLivArea + Neighborhood + GrLivArea*Neighborhood, data = a1_set)
summary(fit)
#plot(fit)
#plot(a1_set$SalePrice, a1_set$GrLivArea)

#put the plots here
```

@@

```{r echo = TRUE, results = 'hide'}
fit_transform <- lm(log(SalePrice) ~ log(GrLivArea) + Neighborhood + log(GrLivArea)*Neighborhood, data = a1_set)
summary(fit_transform)
confint(fit_transform)
#plot(log(a1_set$GrLivArea), log(a1_set$SalePrice))
#plot(fit_transform)

#put the plots here
```

@@

[picture of R table]

@@

```{r}
transformed_numeric = numeric_housing
selected = subset(transformed_numeric, select = -c(LotFrontage, MasVnrArea, GarageYrBlt))
corrplot(cor(selected), tl.cex = 0.5)
```
From this plot, we can eyeball the selection of highly correlated values, including \textit{"OverallQual, TotRmsAbvGrd, GarageCars, TotalBsmtSF, X1stFlrSF, X2ndFlrSF, FullBath"}. By removing these variables from the dataset, we reduce the colinearity of the set and allow for a better feature selection process.

@@

@@

Results Table:

\begin{center}
\begin{tabular}{ |c|c|c|c| } 
 \hline
 Predictive Models & Adjusted R2 & CV PRESS & Kaggle Score \\ 
 \hline
 Forward & 0.8805 & 28.09 & 0.20061 \\
 \hline
 Backward & 0.8812 & 28.12 & 0.20061 \\
 \hline
 Stepwise & 0.8805 & 28.22 & 0.1814 \\
 \hline
 Custom & 
\end{tabular}
\end{center}

@@

>>>>

---
title: "Stats Project 2"
author: 
  - Abhay Sharma | AbhaySh4r.github.io
  - Carolina Craus | carolinacraus.github.io
output:
  pdf_document:
    extra_dependencies: ['graphicx']
  html_notebook: default
---
```{r include=FALSE}
library(tidyverse)
library(plotly)
library(car)
library(caret)
library(olsrr)
library(corrplot)
```
\graphicspath{ {./images/} }

@@

\\

\textbf{Rshiny APP: https://carolinacraus.shinyapps.io/HousingApp/}

@@

For this project, we are provided a housing data set from the website Kaggle that contains a large collection of parameters that are likely to be correlated with the Sale Price of their respective houses. This dataset was built and provided by Century 21 Ames. Within this data set we find 81 parameters and 1460 rows, some of which must ultimately be used to build a predictive model to extrapolate housing prices given a certain set of these parameters. Key variables to take note are \textit{GrLIvArea} which is defined as \textit{Above ground living area}; \textit{Neighborhood}, which is the region that each house is from, and \textit{SalePrice}, which is the price that the house was sold for.

@@

\subsection{Model Generation and Analysis}
To generate an initial model, we will look at a rudimentary approach to linear regression to see what our data does. Generating this model, seen in the Appendix results in the following residual plot.

```{r, echo=FALSE,fig.cap="Residuals for standard linear model"}
knitr::include_graphics("./images/residuals_fit.png")
```

@@

```{r, echo=FALSE,fig.cap="Residuals for transformed linear model"}
knitr::include_graphics("./images/residuals_fit_transform_image.png")
```

@@

```{r, echo=FALSE,fig.cap="Output Paramters of Model Summary", out.width= "60%",fig.align='center'}
knitr::include_graphics("./images/transform_paramters.png")
```

@@

```{r, echo=FALSE,fig.cap="Correlation Matrix"}
knitr::include_graphics("./images/numerical_correlation.png")
```

From this plot, we can eyeball the selection of highly correlated values, including \textit{"OverallQual, TotRmsAbvGrd, GarageCars, TotalBsmtSF, X1stFlrSF, X2ndFlrSF, FullBath"}. By removing these variables from the dataset, we reduce the colinearity of the set and allow for a better feature selection process.

@@

```{r, echo=FALSE,fig.cap="Residuals for backwards elimination model"}
knitr::include_graphics("./images/residuals_backt1.png")
```

@@

\subsection{Comparing Competing Models}

The following is the Results Table, from which we can analyze core components of each selection model:

\begin{center}
\begin{tabular}{ |c|c|c|c| } 
 \hline
 Predictive Models & Adjusted R2 & CV PRESS & Kaggle Score \\ 
 \hline
 Forward & 0.8805 & 28.09 & 0.20061 \\
 \hline
 Backward & 0.8812 & 28.12 & 0.20061 \\
 \hline
 Stepwise & 0.8805 & 28.22 & 0.1814 \\
 \hline
 Custom & 0.846 & 38.28 & 0.20293 \\
 \hline
\end{tabular}
\end{center}

For the models, we restricted paramter selection to only numeric and limited coorelated parameters. This initial approach led to competitive and interesting Kaggle scores. However, our custom approach intended on adding categorical variables to identify more interactions. Our initial approach seemed to lower our Kaggle score, so we didn't pursue that specific model more than what we had.

Ideally we believe that a proper execution of adding in categorical variables to the selection pool should result in a decreased Kaggle Score.

  For Analysis 2, we were required to create 4 prediction models that, that atleast somewhat accurately predicted the sale price of houses sold within the Ames Iowa region. To use the data, we had to address issues of outliers and multicollinearity within the dataset, and narrowed our functional dataset to only numerical values. Using forwards selection, backwards elimination, stepwise selection, and our own custom selected parameter approach, we generated mildly competitive prediction results with Kaggle scores of \textbf{0.20061, 0.20061, 0.1814, 0.20293}, respectively. For the custom approach, although we did not see large gains, we believe that the approach of introducing categorical variables can be successful and should be studied more, for a more thorough model.

\newpage

@@

## Custom Model: # Change categorical values to numeric
housing2 = housing
housing2$Neighborhood <- as.numeric(factor(housing2$Neighborhood,
                                                        levels = c("Blmngtn", "Blueste", "BrDale"),
                                                        labels = c(2,1,0), ordered = TRUE))
housing2$LotConfig <- as.numeric(factor(housing2$LotConfig,
                                                        levels = c("Inside", "Corner", "CulDSac", "FR2", "FR3"),
                                                        labels = c(4,3,2,1,0), ordered = TRUE))
housing2$ExterQual <- as.numeric(factor(housing2$ExterQual,
                                                        levels = c("Ex", "Gd", "TA", "Fa", "Po"),
                                                        labels = c(4, 3, 2,1,0), ordered = TRUE))
housing2$BsmtQual <- as.numeric(factor(housing2$BsmtQual,
                                                        levels = c("Ex", "Gd", "TA", "Fa", "Po"),
                                                        labels = c(4, 3, 2,1,0), ordered = TRUE))

normal_housing = housing 
normal_set = normal_housing %>% select(SaleCondition) %>% filter( SaleCondition == c("Normal"))
normal_housing = normal_housing %>% select_if(is.numeric)
normal_housing$SalePrice = log(normal_housing$SalePrice)
normal_housing$GrLivArea = log(normal_housing$GrLivArea)

t2 = subset(normal_housing, select = -c(LotFrontage, MasVnrArea, GarageYrBlt, OverallQual, TotRmsAbvGrd, GarageCars, TotalBsmtSF, X1stFlrSF, X2ndFlrSF, FullBath )) 
#removal of largely correlated values to reduce multicollinearity (VIF /approx 1) and actually allow back-elimination to run,
#it wasn't running without a ton of errors without this. 
corrplot(cor(t2)) #visualize correlation to identify which plots are correlated. After elimiation, it should be fairly sparsely correlated. 
t2fit = lm(SalePrice ~. , data = t2)
backt2 = ols_step_backward_p(t2fit, prem = 0.05, details = FALSE)
forwardt2 = ols_step_forward_p(t2fit, penter = 0.05, details = FALSE)               #the p-values for these should be fiddled with, 
stepwiset2 = ols_step_both_p(t2fit, pent = 0.05, prem =0.01, details = FALSE)
forwardt2
backt2
stepwiset2
#mystery4thsetWIP
#---------------------------------Summary Stats of the produced RL sets
summary(backt2$model)
vif(backt2$model)
ModelRMSE(backt2$model$residuals)
CVPress(backt2$model)
#---------------------------------
summary(forwardt2$model)
vif(forwardt2$model)
ModelRMSE(forwardt2$model$residuals)
CVPress(forwardt2$model)
#---------------------------------
summary(stepwiset2$model)
vif(stepwiset2$model)
ModelRMSE(stepwiset2$model$residuals)
CVPress(stepwiset2$model)
```
